---
title: 'Lab 5: times series and a map'
author: "Cori Lopazanski"
date: "2/6/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

### 1. Attach packages

```{r}
library(tidyverse)
library(janitor)
library(lubridate)
library(here)
library(paletteer)
library(tsibble)
library(fable) # time series forecasting
library(fabletools)
library(feasts) # time series exploration
library(forecast)
library(sf)
library(tmap)
library(mapview)

```

### Monthly US energy consumption for different types of renewables

```{r}
us_renew <- read_csv(here::here("data", "renewables_cons_prod.csv")) %>% 
  clean_names()

```

- Make description all lowercase
- Only keep observation where 'description' variable contains 'consumption'
- Remove observations where 'description' variable contains 'total'

```{r}
renew_clean <- us_renew %>% 
  mutate(description = str_to_lower(description)) %>% 
  filter(str_detect(description, pattern = "consumption")) %>% 
  filter(!str_detect(description, pattern = "total"))

```

### Convert 'yyyymm' column to a date

```{r}
renew_date <- renew_clean %>% 
  mutate(yr_mo_day = lubridate::parse_date_time(yyyymm, "ym")) %>% 
  mutate(month_sep = yearmonth(yr_mo_day)) %>% 
  mutate(value = as.numeric(value)) %>% 
  drop_na(month_sep, value)
  
# Make version where month and year are in separate columns
renew_parsed <- renew_date %>% 
  mutate(month = month(yr_mo_day, label = TRUE),
         year = year(yr_mo_day))


```


### Plot

```{r}
renew_gg <- ggplot(data = renew_date, aes(x = month_sep, y = value)) +
  geom_line(aes(color = description))

renew_gg 
```

Updating colors with paleteer palettes:
`paleteer` package shows discrete packages - want to find one look in View(palettes_d_names)

```{r}
renew_gg + 
  scale_color_paletteer_d("calecopal::figmtn")
```

## Convert to a tsibble

Convert to a time-series-enabled dataframe for friendlier working with fable and forecast

```{r}
renew_ts <- as_tsibble(renew_parsed, key = description, index = month_sep)
```

Look at the time series data in a few different ways:
```{r}
renew_ts %>% autoplot(value)
renew_ts %>% gg_subseries(value)
renew_ts %>% gg_season(value)

# Make the season plot in ggplot instead (sometimes it doesn't work)

ggplot(data = renew_parsed, aes(x = month, y = value, group = year))+
  geom_line(aes(color = year))+
  facet_wrap(~description, ncol = 1, scales = "free",
             strip.position = "right")
  
```

Just look at hydroelectric energy consumption

```{r}
hydro_ts <- renew_ts %>% 
  filter(description == "hydroelectric power consumption") 

hydro_ts %>% autoplot(value)
hydro_ts %>% gg_subseries(value)
hydro_ts %>% gg_season(value)
```


What if I want quarterly average consumption for hydro?

- Can index over whatever groupings you want

```{r}
hydro_quarterly <- hydro_ts %>% 
  index_by(year_qu = ~(yearquarter(.))) %>% 
  summarize(avg_consumption = mean(value))
```


### Decompose that hydro_ts using functions from feasts and fable

```{r}
dcmp <- hydro_ts %>% 
  model(STL(value ~ season(window = 5)))
  
components(dcmp) %>% autoplot()
hist(components(dcmp)$remainder) # pretty normally distributed

  
```


Examine ACF (autocorrelation function)
```{r}
hydro_ts %>% 
  ACF(value) %>% autoplot()
```

### DANGER DANGER 

Whenever you do some modeling, you need to make sure you're doing your research and reading and doing the right thing... but here's a preview 

```{r}
hydro_model <- hydro_ts %>% 
  model(
    ARIMA(value)
  ) %>% 
  fabletools::forecast(h = "4 years")

hydro_model %>% autoplot(filter(hydro_ts, year(month_sep) > 2010))

```


### Make a world map!

```{r}
world <- read_sf(dsn = here("data", "TM_WORLD_BORDERS_SIMPL-0.3-1"), 
                 layer = "TM_WORLD_BORDERS_SIMPL-0.3")

mapview(world)

```







